---
title: "PSTAT 126 HW 1"
author: "Nathan Wu, 6134910"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: yes
---

```{r setup, echo=FALSE}
# set global chunk options: images will be 6x4 inches
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center', fig.width=6, fig.height=4)
options(digits = 4)
```
\newpage

```{r packages}
 library(alr4)
 library(faraway)
```

# 1. Linear regression model of `ppgdp` on `fertility`.

Import the data set *UN11* in *alr4* package, extract the data *ppgdp* and *fertility*
```{r}
data('UN11')
x = UN11$fertility
y = UN11$ppgdp
```

## (a) Scatterplot of `ppgdp` on `fertility`
Below is the scatterplot of *ppgdp* against *fertility*
```{r}
plot(x, y, xlab = "Fertility (Children/Women)", ylab = "Per Capita GDP ($)")
```
From the above graph, we can't see a very clear linear relationship between these two data
Now, calculate the natural logrithms of both variables 
```{r}
logx = log(x)
logy = log(y)
```

Replacing both variables by their natural logrithms, and draw another scatterplot

## (b) Scatterplot of log(`ppgdp`) against log(`fertility`)
```{r}
plot(logx, logy, xlab = "log Fertility", ylab = "log Per Capita GDP")
```

Now we can see a plausible linear trend that we use for summary. 

## (c) compute lse for slope and intercept 
Now we want to find the LSE
```{r}
xbar = mean(logx)
ybar = mean(logy)
Sxx = sum((logx-xbar)^2)
Syy = sum((logy-ybar)^2)
Sxy = sum((logx-xbar)*(logy-ybar))
b1 = Sxy/Sxx
b0 = ybar - b1*xbar
plot(logx, logy, xlab = "log Fertility", ylab = "log Per Capita GDP")
abline(b0,b1)
b1
b0
```

# 2. Predict `lpsa` with `lcavol`
First, import the data set *prostate*
```{r}
data('prostate')
y = prostate$lpsa
x = prostate$lcavol
```
After extracting the data from the data set, we want to draw a scatter plot to discover the relationship between these two data
## (a)
```{r}
plot(x , y, xlab = "log cancer volume", ylab = "log prostate specific antigen") 
```
We can see there is a quite clear postive linear relationship between these two data

## (b)
```{r}
xbar = mean(x)
ybar = mean(y)
Sxx = sum((x-xbar)^2)
Syy = sum((y-ybar)^2)
Sxy = sum((x-xbar)*(y-ybar))
b1 = Sxy/Sxx
b0 = ybar - b1*xbar
plot(x , y, xlab = "log cancer volume", ylab = "log prostate specific antigen") 
abline(b0,b1)
xbar
ybar
Sxx
Sxy
Syy
```

## (c) Obtain the estimate of σ2 and find the estimated standard errors of β􏰀0 and β􏰀1
```{r}
yhat = b0 + b1*x
e = y - yhat
n = length(x)
sse = sum((y-yhat)^2)
mse = sse/(n-2)
se_b1 = sqrt(mse/Sxx)
se_b0 = sqrt(mse* ((1/n) + ((xbar^2)/Sxx) ) )

mse
se_b0
se_b1
```

## (d) Find the estimated covariance between β0 and β1.
```{r}
-mse*xbar/Sxx
```

## (e) Carry out t-tests for the two null hypotheses β0 = 0 and β1 = 0, reporting the value of the test statistic and a p-value in each case.
```{r}
test_b0 = b0/se_b0
test_b1 = b1/se_b1


p_b0 = 2*pt(q = test_b0, df = n-2, lower.tail = F)
p_b1 = 2*pt(q = test_b1, df = n-2, lower.tail = F)

test_b0
test_b1
p_b0
p_b1
```
According to the result of p-value of $\beta_1$, we are 95% sure that there is a positive linear relationship between x and Y. 


## (f) Use the R function lm() to fit the regressions of lpsa on lcavol.
```{r}
fit = lm(y~x)
summary(fit)
```

